---
title: "Evaluate evapotranspiration simulation with rwrfhydro"
author: "Aubrey Dugger"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# Background
We are using WRF-Hydro to predict streamflow for Fourmile Creek at the Orodell USGS gage for the 2013 snowmelt period. We ran WRF-Hydro with NoahMP as the LSM for a 3-year spinup period and then did an hourly run for 1 month starting May 1, 2013. We want to evaluate the ET fluxes at a local Ameriflux tower site.

Load the rwrfhydro package. 
```{r}
library("rwrfhydro")
```

Set a data path to the Fourmile Creek test case.
```{r}
dataPath <- '/Users/adugger/WRF_Hydro/test_cases/Fourmile_Creek'
```


# Import modelled datasets

Import LSM water fluxes at the cell where the Ameriflux tower is located. Since we do NOT want a basin-wide average, but instead a time series of values for a specified cell, we will use the GetMultiNcdf tool directly. First, we do a bit of setup.

Setup the cell you want to extract and the stat you want to calculate. We are only running a single cell (i=2, j=3, time=1) so we specify the same start and end cell and use "CalcMeanMinrm" as the stat. "CalcMeanMinrm" is a utility function that calculates a mean with an enforced minimum valid value (-1e+30). This is useful to remove intended "no data" values that were not caught by NetCDF checks.

```{r}
niw <- list(start=c(2,3,1), end=c(2,3,1), stat='CalcMeanMinrm')
```

Now setup which LDASOUT variables we want to bring in. We will be working with the accumulated ET fluxes at our daily time step.
```{r}
variableNames <- c('ACCECAN','ACCEDIR','ACCETRAN','ACCPRCP') 
ldasoutVars <- as.list( variableNames ) 
names(ldasoutVars) <- variableNames
ldasoutVariableList <- list( ldasout = ldasoutVars )
```

Match the variables to the desired cell and stat (here we are using the same for all variables).
```{r}
ldasoutInd <- list( niw, niw, niw, niw )
names(ldasoutInd) <- names(ldasoutVars)
ldasoutIndexList <- list( ldasout = ldasoutInd )
```

Generate the list of filenames for the GetMultiNcdf tool by providing the pathname to the model OUTPUT directory and the filename pattern to match (in this case we want the LDASOUT files).
```{r}
ldasoutFilesList <- list( ldasout = list.files(path=paste0(dataPath, '/OUTPUT_ALLRT'), 
     pattern=glob2rx('*LDASOUT*'), full.names=TRUE))
```

Run the tool. To speed things up, we will run this over 8 cores using the doMC package.
```{r}
library(doMC)
registerDoMC(8)
ldasoutDf <- GetMultiNcdf(ind=ldasoutIndexList, var=ldasoutVariableList, files=ldasoutFilesList, parallel=T )
```

Now we can "flatten" the output to make it work better with certain tools.
```{r}
modLdasout.allrt.niw <- ReshapeMultiNcdf(ldasoutDf)
rm(ldasoutDf)
```

Some of the LDASOUT fields we exported are accumulated totals, so we need to break them out into individual (per time step) fluxes.
```{r}
modLdasout.allrt.niw <- CalcNoahmpFluxes(modLdasout.allrt.niw)
```

We want to examine total ET, so we need to calculate a sum column for all 3 evaporation and transpiration fluxes (in units of "mm"). We then use the density of water (1mm H<sub>2</sub>O * m<sup>2</sup> = 1 kg), the latent heat of vaporization (assuming evaporation and not sublimation during this summer period, LHVAP = 2.5106e+6 J/kg), and the time step (1 day = 86400 sec) to calculate total daily latent heat flux in W/m<sup>2</sup>.
```{r}
modLdasout.allrt.niw$ET_mm <- with(modLdasout.allrt.niw, DEL_ACCECAN + DEL_ACCETRAN + DEL_ACCEDIR)
modLdasout.allrt.niw$LE <- with(modLdasout.allrt.niw, ET_mm * 2.5106e+6 / 86400)
```

# Import observed datasets

Import Ameriflux station data for Niwot Ridge (Level 2 Standardized). We specify "America/Denver" as the time zone (Ameriflux timestamps are in local time).
```{r}
obsFlux30min.usnr1 <- ReadAmerifluxCSV(paste0(dataPath, '/OBS/AMF_USNR1_2013_L2_GF_V008.csv'), "America/Denver")
```

Our LDASOUT time step was 1 day, so we need to aggregate the observations to a UTC daily time step for comparison.
```{r}
obsFlux1d.usnr1 <- aggregate(obsFlux30min.usnr1[names(obsFlux30min.usnr1) != c("POSIXct")], 
     by = list(as.POSIXct(round(as.POSIXct(format(obsFlux30min.usnr1$POSIXct, tz="UTC"), tz="UTC"), "days"))), 
     CalcMeanNarm)
names(obsFlux1d.usnr1)[1] <- "POSIXct"
```


# Plot & compare the time series of ET fluxes

```{r compFluxes, fig.width = 12, fig.height = 6, out.width='700', out.height='350'}
PlotFluxCompare(obsFlux1d.usnr1, "LE", modLdasout.allrt.niw, "LE", 
     labelObs="Observed Niwot Ameriflux", 
     labelMod1="All Routing", title="ET: Niwot Ridge")
```


# Review model performance statistics

```{r, results='hide'}
CalcModPerf(modLdasout.allrt.niw, obsFlux1d.usnr1, flxCol.mod="LE", flxCol.obs="LE")
```

```{r, results = "asis", echo=FALSE}
library(pander)
pander::pandoc.table(CalcModPerf(modLdasout.allrt.niw, obsFlux1d.usnr1, flxCol.mod="LE", flxCol.obs="LE"),split.table=Inf)
```
