---
title: "Evaluate streamflow simulation with rwrfhydro"
author: "Aubrey Dugger and James McCreight"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` setion of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))



# Background: 
ßWe are using WRF-Hydro to predict streamflow for Fourmile Creek at the Orodell USGS 

gage for the 2013 water year (excluding the September 2013 flood event). We ran WRF-Hydro with 

NoahMP as the LSM for a 3-year spinup period and then did an hourly run for 11 months starting October 

1, 2012. We want to evaluate model performance in estimating streamflow with and without overland, 

subsurface, and groundwater routing active.

############################################

## Import modelled and observed datasets

# Model 1: Only channel routing turned on (hourly model run)

modStrh.chrt.fc <- ReadFrxstPts("OUTPUT_FOURMILE_CHRT/frxst_pts_out.txt")

# Model 2: All WRF-Hydro routing options turned on (hourly model run)

modStrh.allrt.fc <- ReadFrxstPts("OUTPUT_FOURMILE_ALLRT/frxst_pts_out.txt")

# USGS gage observed data @ 5-minute intervals

obsStr5min.fc <- ReadUsgsGage("OBS/5min_str_06727500_110401_140810.txt")

############################################

## Plot & compare hydrographs for the full model run

PlotFluxCompare(obsStr5min.fc, "q_cms", modStrh.chrt.fc, "q_cms", strDf.mod2=modStrh.allrt.fc, 

strCol.mod2="q_cms", labelObs="Observed Fourmile Creek at Orodell", labelMod1="Channel 

Routing Only", labelMod2="All Routing", title="Streamflow: Fourmile Creek")

# Now limit the plot to the May through June snowmelt period. Note that the reported stats are also 

updated to the new time period.

PlotFluxCompare(obsStr5min.fc, "q_cms", modStrh.chrt.fc, "q_cms", strDf.mod2=modStrh.allrt.fc, 

strCol.mod2="q_cms", labelObs="Observed Fourmile Creek at Orodell", labelMod1="Channel 

Routing Only", labelMod2="All Routing", title="Streamflow: Fourmile Creek", 

stdate=as.POSIXct("2013-05-01 00:00:00", format="%Y-%m-%d %H:%M:%S", tz="UTC"), 

enddate=as.POSIXct("2013-06-30 00:00:00", format="%Y-%m-%d %H:%M:%S", tz="UTC"))

############################################

## Review flow duration curves

# Calculate percent exceedances for flow duration curves

obsStr5min.fc <- CalcFdc(obsStr5min.fc)

modStrh.chrt.fc <- CalcFdc(modStrh.chrt.fc)

modStrh.allrt.fc <- CalcFdc(modStrh.allrt.fc)

# Compare how the models are doing predicting flow values that will be exceeded 20% of the time

# First, calculate the fitted spline functions

fdc.obsStr5min.fc <- CalcFdcSpline(obsStr5min.fc)

fdc.modStrh.chrt.fc <- CalcFdcSpline(modStrh.chrt.fc)

fdc.modStrh.allrt.fc <- CalcFdcSpline(modStrh.allrt.fc)

# Then, evaluate at the 20% exceedance percentage (high flows)

fdc.obsStr5min.fc(0.2)

fdc.modStrh.chrt.fc(0.2)

fdc.modStrh.allrt.fc(0.2)

# Now try the 80% exceedance percentage (low flows)

fdc.obsStr5min.fc(0.8)

fdc.modStrh.chrt.fc(0.8)

fdc.modStrh.allrt.fc(0.8)

# Plot flow duration curves for a more complete picture

PlotFdcCompare(obsStr5min.fc, "q_cms", modStrh.chrt.fc, "q_cms", strDf.mod2=modStrh.allrt.fc, 

strCol.mod2="q_cms", labelObs="Observed Fourmile Creek", labelMod1="Channel Routing 

Only", labelMod2="All Routing")

############################################

## Review model performance statistics

# Calculate model performance stats

CalcModPerf(modStrh.allrt.fc, obsStr5min.fc)

CalcModPerf(modStrh.chrt.fc, obsStr5min.fc)