---
title: "Evaluate streamflow simulation with rwrfhydro"
author: "Aubrey Dugger and James McCreight"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# Background
We are using WRF-Hydro to predict streamflow for Fourmile Creek at the Orodell USGS gage for the 2013 snowmelt period. We ran WRF-Hydro with NoahMP as the LSM for a 3-year spinup period and then did an hourly run for 1 month starting May 1, 2013. We want to evaluate model performance in estimating streamflow with and without overland, subsurface, and groundwater routing active.

Load the rwrfhydro package. 
```{r}
library("rwrfhydro")
```

Set a data path to the Fourmile Creek test case.
```{r}
dataPath <- '/Users/adugger/WRF_Hydro/test_cases/Fourmile_Creek'
```


# Import modelled and observed datasets

Model 1: Only channel routing turned on (hourly model run).
```{r}
modStrh.chrt.fc <- ReadFrxstPts(paste0(dataPath, '/OUTPUT_CHRT/frxst_pts_out.txt'))
```

Model 2: All WRF-Hydro routing options turned on (hourly model run).
```{r}
modStrh.allrt.fc <- ReadFrxstPts(paste0(dataPath, '/OUTPUT_ALLRT/frxst_pts_out.txt'))
```

USGS gage observed data at 5-minute intervals.
```{r}
obsStr5min.fc <- ReadUsgsGage(paste0(dataPath, '/OBS/5min_str_06727500_110401_140810.txt'))
```


# Plot hydrographs 

Compare hydrographs for the full model run.
```{r compHydrographs, fig.width = 12, fig.height = 6, out.width='700', out.height='350'}
PlotFluxCompare(obsStr5min.fc, "q_cms", modStrh.chrt.fc, "q_cms", strDf.mod2=modStrh.allrt.fc, 
                strCol.mod2="q_cms", labelObs="Observed Fourmile Creek at Orodell", 
	            	labelMod1="Channel Routing Only", labelMod2="All Routing", 
		            title="Streamflow: Fourmile Creek")
```

Now limit the plot to the late May recession period only. The reported stats are updated to the new time period. (Note that the R warning is innocuous because the subset adjusts for timezone, so it's ok that the timezones don't match.)
```{r compHydrographsSnow, fig.width = 12, fig.height = 6, out.width='700', out.height='350'}
PlotFluxCompare(obsStr5min.fc, "q_cms", modStrh.chrt.fc, "q_cms", strDf.mod2=modStrh.allrt.fc, 
                strCol.mod2="q_cms", labelObs="Observed Fourmile Creek at Orodell", 
            		labelMod1="Channel Routing Only", labelMod2="All Routing", 
		            title="Streamflow: Fourmile Creek", 
                stdate=as.POSIXct("2013-06-01 00:00:00", format="%Y-%m-%d %H:%M:%S", tz="UTC"), 
                enddate=as.POSIXct("2013-06-30 00:00:00", format="%Y-%m-%d %H:%M:%S", tz="UTC"))
```

# Review flow duration curves

Calculate percent exceedances for flow duration curves.
```{r}
obsStr5min.fc <- CalcFdc(obsStr5min.fc)
modStrh.chrt.fc <- CalcFdc(modStrh.chrt.fc)
modStrh.allrt.fc <- CalcFdc(modStrh.allrt.fc)
```

Compare how the models are doing predicting flow values that will be exceeded 20% of the time.
First, calculate the fitted spline functions.
```{r}
fdc.obsStr5min.fc <- CalcFdcSpline(obsStr5min.fc)
fdc.modStrh.chrt.fc <- CalcFdcSpline(modStrh.chrt.fc)
fdc.modStrh.allrt.fc <- CalcFdcSpline(modStrh.allrt.fc)
```

Then, evaluate at the 20% exceedance percentage (high flows).
```{r, results='hold'}
fdc.obsStr5min.fc(0.2)
fdc.modStrh.chrt.fc(0.2)
fdc.modStrh.allrt.fc(0.2)
```

Now try the 80% exceedance percentage (low flows).
```{r, results='hold'}
fdc.obsStr5min.fc(0.8)
fdc.modStrh.chrt.fc(0.8)
fdc.modStrh.allrt.fc(0.8)
```

Plot flow duration curves for a more complete picture.
```{r flowDurationCurves, fig.width = 12, fig.height = 6, out.width='700', out.height='350'}
PlotFdcCompare(obsStr5min.fc, "q_cms", modStrh.chrt.fc, "q_cms", strDf.mod2=modStrh.allrt.fc, 
               strCol.mod2="q_cms", labelObs="Observed Fourmile Creek", 
		           labelMod1="Channel Routing Only", labelMod2="All Routing")
```


# Review model performance statistics

Calculate model performance stats (special formatting comands hidden).
```{r, results='hide'}
CalcModPerf(modStrh.allrt.fc, obsStr5min.fc)
```

```{r, , results = "asis", echo=FALSE}
library(pander)
pander::pandoc.table(CalcModPerf(modStrh.allrt.fc, obsStr5min.fc),split.table=Inf)
```

```{r, results='hide'}
CalcModPerf(modStrh.chrt.fc, obsStr5min.fc)
```

```{r, , results = "asis", echo=FALSE}
pander::pandoc.table(CalcModPerf(modStrh.chrt.fc, obsStr5min.fc), split.table=Inf)
```

Help on CalcModPerf gives details on the individual statistics returned.
```{r, results='hide'}
<<background='#FF0000'>>
help(CalcModPerf)
```

<div style="border:1px solid; border-radius: 25px; padding: 12px 25px;">
```{r, echo=FALSE}
<<background='#FF0000'>>
library(printr)
help(CalcModPerf)
```
</div>

Calculate flow duration curve performance statistics.
```{r, results='hide'}
CalcFdcPerf(modStrh.allrt.fc, obsStr5min.fc)
```

```{r, , results = "asis", echo=FALSE}
pander::pandoc.table(CalcFdcPerf(modStrh.allrt.fc, obsStr5min.fc),split.table=Inf)
```

Again, help on CalcFdcPerf gives details on the individual statistics returned.
```{r, results='hide' }
help(CalcFdcPerf)
```
<div style="border:1px solid; border-radius: 25px; padding: 12px 25px;">
```{r, echo=FALSE }
help(CalcFdcPerf)
```
</div>